---
title: "Practical Machine Learning Assessment"
author: "Rowen Remis R. Iral"
date: "Friday, August 22, 2014"
output: html_document
---

#Practical Machine Learning Assessment
### Applying machine learning using R
  Jawbone Up, Nike FuelBand, and Fitbit can now collect vast amount of data using devices.  With these devices we can now quantify alot of data.
  
### Set your Working Directory
```{r}
# setwd("E:/Other Business/Data Scientist/Data Science Track/Practical Machine Learning/assessment_project/PracticalMachineLearning_Assessment/")
```

###Data Sources
The training data for this project are available here: 
  
  https://d396qusza40orc.cloudfront.net/predmachlearn/pml-training.csv

The test data are available here: 
  
  https://d396qusza40orc.cloudfront.net/predmachlearn/pml-testing.csv

http://groupware.les.inf.puc-rio.br/har - Weight Lifting Exercise data

###knitr Settings
```{r}
library(knitr)
opts_chunk$set(echo = TRUE, results = 'hold')
```

###Load Library
```{r}
library(ElemStatLearn)
library(caret)
library(rpart)
library(randomForest)
set.seed(123)
```

###Data
```{r}
library(ggplot2)
pml_train <- read.csv("pml-training.csv", header=TRUE, sep=",", na.strings=c("NA",""))

pml_train <- pml_train[,-1] #Remove Column ID row

#pml_train$interval <- factor(pml_train$interval)
#pml_train$date <- as.Date(pml_train$date, format="%Y-%m-%d")

pml_test <- read.csv("pml-testing.csv", header=TRUE, sep=",", na.strings=c("NA",""))

pml_test <- pml_test[,-1]

#pml_test$interval <- factor(pml_test$interval)
#pml_test$date <- as.Date(pml_test$date, format="%Y-%m-%d")
```

###Counting Training data
```{r}
dim(pml_train)
```
There are 19622 rows and 160 variables in pml_train data.

###Counting Testing data
```{r}
dim(pml_test)
```
There are 20 rows and 160 columns in the pml_test data.

###Get Variable names in pml_train
```{r}
names(pml_train)
```

###Get Variable names in pml_test
```{r}
names(pml_test)
```


###Modeling Training Data
```{r}
library(caret)
inTrain = createDataPartition(pml_train$classe, p=0.60, list=FALSE)
training = pml_train[inTrain,]
validating = pml_train[-inTrain,]
```


###Get number of columns with less than 60% of data
```{r}
sum((colSums(!is.na(training[,-ncol(training)])) < 0.6*nrow(training)))
```

###Remove columns that doesn't have data
```{r}
Keep <- c((colSums(!is.na(training[,-ncol(training)])) >= 0.6*nrow(training)))
training   <-  training[,Keep]
validating <- validating[,Keep]
```

###Random Forest Modeling
```{r}
library(randomForest)
model <- randomForest(classe~.,data=training)
print(model)
```

###Evaluate Model
Get the important data
```{r}
importance(model)
```

###Use Confusion Matrix for Evaluation
```{r}
confusionMatrix(predict(model,newdata=validating[,-ncol(validating)]),validating$classe)
```

###Calculate accuracy
The accuracy is: 99.84706 %
```{r}
accuracy<-c(as.numeric(predict(model,newdata=validating[,-ncol(validating)])==validating$classe))
accuracy<-sum(accuracy)*100/nrow(validating)
print(accuracy)
```

###Modeling Test Data
```{r}
pml_test <- pml_test[ , Keep] #Keep the same columns as of testing dataset
pml_test <- pml_test[,-ncol(pml_test)] # Remove the Problem ID
```

###Transform and Coerce Testing Data Set
```{r}
testing <- rbind(training[100, -59] , pml_test) 
row.names(testing) <- c(100, 1:20)
```

###Predict with Test Data
```{r}
predictions <- predict(model,newdata=testing[-1,])
print(predictions)
```

###Submitting Answers
```{r}
pml_write_files = function(x){
  n = length(x)
  for(i in 1:n){
    filename = paste0("problem_id_",i,".txt")
    write.table(x[i],file=filename,quote=FALSE,row.names=FALSE,col.names=FALSE)
  }
}

pml_write_files(predictions)
```

The answers will be checked by the checker, as most random forest predicts with high accuracy, we can see its results.