---
title: "Practical Machine Learning Assessment"
author: "Rowen Remis R. Iral"
date: "Sunday, July 27, 2014"
output: html_document
---

#Practical Machine Learning Assessment
### Applying machine learning using R
  Jawbone Up, Nike FuelBand, and Fitbit can now collect vast amount of data using devices.  With these devices we can now quantify alot of data.

###Data Sources
The training data for this project are available here: 
  
  https://d396qusza40orc.cloudfront.net/predmachlearn/pml-training.csv

The test data are available here: 
  
  https://d396qusza40orc.cloudfront.net/predmachlearn/pml-testing.csv

http://groupware.les.inf.puc-rio.br/har - Weight Lifting Exercise data

###Data
```{r}
library(ggplot2)
training <- read.csv("pml-training.csv")
testing <- read.csv("pml-testing.csv")
```

###Counting Training data
```{r}
dim(training)
```
There are 19622 rows and 160 variables in training data.

###Counting Testing data
```{r}
dim(testing)
```
There are 20 rows and 160 columns in the testing data.

###Get Variable names
```{r}
names(training)
```


###Caret modeling
```{r}
library(caret)
library(rpart)
modFit <- train(classe ~ ., data=training, method="rpart")
print(modFit$finalModel)
```

###Plot Tree
```{r}
plot(modFit$finalModel, uniform = TRUE, main="Classe Tree")
text(modFit$finalModel, use.n=TRUE, all=TRUE, cex=.8)
```

###Prettier Plot
```{r}
library(rattle)
fancyRpartPlot(modFit$finalModel)
```

###Predicting new values
```{r}
predict(modFit,newdata=testing)
```


